Tidy data structure to support exploration and modeling of temporal-context data

### 2018-08-17 main.Rmd
rewrite hts

On the other hand, those who are engaged in longitudinal design, are primarily interested in explaining variations among individuals in change process and making inference about the underlying population. Longitudinal data or panel data typically refers to fewer measurements (small $T$) over a large number of individuals (large $N$). It often occurs to longitudinal design that some individuals are not observed at certain points in time, resulting in an unbalanced panel. This data issue leads representing longitudinal data much more commonly in **long format** where each row corresponds to an observational unit per time and each column contains measurements across the individuals for all points in time. Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it in that wide format. Another appealing feature for a long format is data structured in a concise and semantic manner with reference to observations and variables. This subsequently supports perceiving time index as an explicit variable instead of an implicit attribute. In general, fewer restrictions are made if representing temporal-context data in long format.

Longitudinal data representation has been implemented in Stata's time series module [@Stata] and R package **plm** [@plm2008]. The underpinning data structure is a two-dimensional column-homogeneous array. Specifying a longitudinal data set from a generic array needs explicitly declaring time and individuals (or panel variable in Stata's `tsset` command). Therefore, each row of the data can be identified by a specific individual and time point. Individuals, however, can be only declared through a single variable, not multiple.

@@

Temporal data can often be aggregated in a manner that exhibits a nested or crossed structure, also known as hierarchical or grouped time series [@fpp]. For example, in a manufacturing setting, a company can add up every store’s sales by region, by state and by country, which gives a strictly hierarchical time series; alternatively they can group the sales for each product together based on common attributes such as store, brand, price range and so forth, which leads to a non-hierarchical structure---a grouped time series. Additionally, hierarchical structure can be derived from statistical models that classifies the time series. The R package **hts** [@R-hts] is the implementation of this type of time series. But it is frustrating to create and work with such data objects due to non-extensible and non-modified. Variational data aggregation (some groups no need aggregation). [R]

>>>>

On the other hand, those who are engaged in longitudinal design, are primarily interested in explaining variations among individuals in change process and making inference about the underlying population. Longitudinal data or panel data typically refers to fewer measurements (small $T$) over a large number of individuals (large $N$). It often occurs to longitudinal design that some individuals are not observed at certain points in time, resulting in an unbalanced panel. This data issue leads representing longitudinal data much more commonly in **long format** where each row corresponds to an observational unit per time and each column contains measurements across the individuals for all points in time. Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it in that wide format. Another appealing feature for a long format is data structured in a concise and semantic manner with reference to observations and variables. This supports perceiving time index as an explicit variable instead of an implicit attribute, and subsequently opens the door to manipulating time such as extracting time components. In general, fewer restrictions are made if representing temporal-context data in long format.

Longitudinal data representation has been implemented in Stata's time series module [@Stata] and R package **plm** [@plm2008]. The underpinning data structure is a two-dimensional column-homogeneous array, as other tabular data. Specifying a longitudinal data set from a generic array needs explicitly declaring time and individuals (or panel variable in Stata's `tsset` command). Therefore, each row of the data can be identified by a specific individual and time point. Individuals, however, can be only declared through a single variable, not multiple.

@@

A collection of time series can often be arranged along multivariate data of nested and crossed factors, which is also known as hierarchical and grouped time series [@fpp]. For example, from an operational point of view, a manufacturer can add up every store’s sales by region, by state, and by country along a geographical tree; alternatively, one can group the sales for each product together based on common attributes such as store, brand, price range and so forth, resulting in a non-hierarchical structure. The former gives an example of hierarchical time series, and the latter shows grouped time series. Hierarchical time series can also be derived using hierarchical clustering.

The R package **hts** [@R-hts] is the implementation of this type of time series. But it is frustrating to create and work with such data objects due to non-extensible and non-modified. Variational data aggregation (some groups no need aggregation). [R]
